Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-01 paper_jasa.tex
comment on power definition
model notation updated

@@

In the case that observer  $i = 1, ..., K$ is exposed to multiple lineups we can estimate the probability $p_{i \ell}$ for each lineup $\ell = 1, ..., L$, while accommodating both for different abilities of individuals as well as differences in the difficulty of lineups with the help of a logistic regression using random effects:

Let $X_{i\ell} \sim Binom_{1, p_{i\ell}}$ be the random variable, with $X_{i\ell}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
We can model $p_{i\ell}$ with a logistic regression with subject specific random effects as:
\[
\text{logit } p_{i\ell} = X_\ell B + Z u_i,  
\]
where $X_\ell$ is a matrix of fixed effects with parameter vector $B$ corresponding to lineup $\ell$ and $u \sim MVN(0, \sigma_u I)$  is a vector of random variables $i = 1, ..., K$.

@@

We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}. Suppose, we have responses from $N$ independent observers with $u$ identifications of the actual plot. This gives an estimated power of
\begin{equation}\label{power_est} Power(\beta)=\frac{u}{N} \hspace{1cm} 0 \le u \le N \end{equation}

Suppose each of $N$ independent observers gives evaluations on multiple lineup plots and responses are associated with binary random variable $Y_{ij}$. Let $Y_{ij} = 1$, if subject $j$ correctly identifies the test statistic on lineup $i$, and 0 otherwise.
We model $\pi_{ij} = E(Y_{ij})$ as  a mixed effects logistic regression

\begin{equation} 
g(\pi_{ij}) = X_{ij}B  + Z_{ij} \tau_j 
\label{mixed} 
\end{equation}
%
where $\tau_j$ is random effects coefficient vector of length $q$ for subject $j$,  $\tau_j  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{ij}$ is the $i$th row vector of random effects covariates for subject $j$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{ij}$ is the $i$th row vector of the fixed effects covariates for subject $j$ and logit link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$. The covariates could be demographic information of people such as age, gender, education level etc. as well as sample size ($n$), regression parameter ($\beta$) and variance ($\sigma^2$) of model \ref{multi}.

>>>>

\red{there may be a flow in this definition. Due to the discrete nature of binomial (specially $x_{\alpha}$) this definition would only give a conservative lower limit of the actual power. I feel like we should not define power using binomial.}

@@

In the case that observer  $i = 1, ..., K$ is exposed to multiple lineups we can estimate the probability $p_{ \ell i}$ for each lineup $\ell = 1, ..., L$, while accommodating both for different abilities of individuals as well as differences in the difficulty of lineups with the help of a logistic regression using random effects:

Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
We can model $p_{ \ell i}$ with a logistic regression with subject specific random effects as:
\[
\text{logit } p_{\ell i} = X_{\ell i} B +  Z_{\ell i} \tau_i,  
\]
where $X_{\ell i}$ is a vector of fixed effects with parameter vector $B$ corresponding to lineup $\ell$ and $\tau \sim MVN(0, \sigma_{\tau} I)$  is a vector of random variables $i = 1, ..., K$.

\red{This model is elaborately described in equation \ref{mixed}.}

@@

We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}. Suppose, we have responses from $K$ independent observers with $u$ identifications of the actual plot. This gives an estimated power of
\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}

Suppose each of $K$ independent observers gives evaluations on multiple lineup plots and responses are associated with binary random variable $Y_{\ell i}$. Let $Y_{\ell i} = 1$, if subject $i$ correctly identifies the test statistic on lineup $\ell$, and 0 otherwise.
We model $\pi_{\ell i} = E(Y_{\ell i})$ as  a mixed effects logistic regression

\begin{equation} 
g(\pi_{\ell i}) = X_{\ell i}B  + Z_{\ell i} \tau_i 
\label{mixed} 
\end{equation}
%
where $\tau_i$ is random effects coefficient vector of length $q$ for subject $i$,  $\tau_i  \sim  MVN(0,\Sigma)$ with variance covariance matrix $\Sigma$, $Z_{\ell i}$ is the $\ell$th row vector of random effects covariates for subject $i$, $B$ is a vector of coefficient of length $p$, the number of fixed effect covariates being used, $X_{\ell i}$ is the $\ell$th row vector of the fixed effects covariates for subject $i$ and logit link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$. The covariates could be demographic information of people such as age, gender, education level etc. as well as sample size ($n$), regression parameter ($\beta$) and variance ($\sigma^2$) of model \ref{multi}.
